Asteroid project summary

### 2021-11-13 summaries/21-11/21-11.tex
Finished analysis and runge-kutta

\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{bbold}

@@

\newcommand{\unit}[1]{\hat{\mathbf{#1}}}
\newcommand{\parens}[1]{\left( #1 \right)}
\newcommand{\brackets}[1]{\left[ #1 \right]}
\newcommand{\jtd}[1]{{\color{red}\textit{#1}}}
\usetikzlibrary{arrows.meta}

@@

We define a rotation matrix $M$ such that $M\bm{r} = \bm{R}$, where $\bm{r}$ is in the body-fixed frame and $\bm{R}$ is in the inertial frame. We write this matrix in terms of the $z-y-z$ Euler angles, named $\alpha$, $\beta$, and $\gamma$, defined in figure \ref{fig:euler-angles}. These Euler angles, together with the position and momentum of the asteroid, are our dynamical variables.

@@

\subsection{Potential energy}
We start by expressing the gravitational potential energy of the system in its most general form:
\begin{equation}
V = -G\int_\mathcal{A} d^3 r' \rho_m(\bm r')\int_\mathcal{P} d^3 r \rho_M(\bm r) \frac{1}{|\bm{r}-\bm{r'}|}.
\label{eqn:first-pe}
\end{equation}
This integral is computed in the inertial frame. Here, $\bm{\omega}$ is the rotational velocity of the asteroid and $I$ is the inertia tensor, both in the body fixed frame. We also use the asteroid density $\rho_m$ and the central body density $\rho_M$. The symbol $\mathcal{A}$ denotes the volume of the asteroid and $\mathcal{P}$ denotes the volume of the central body. To reduce equation \ref{eqn:first-pe}, we define the unnormalized spherical harmonics $Y_{\ell m}(\theta, \phi) = P_{\ell m}(\cos \theta)e^{im\phi}$, where $P_{\ell m}$ are the associated Legendre Polynomials without the Condon-Shortley phase. The regular and irregular spherical harmonics are then defined as
\begin{equation}
S_{\ell m}(\bm r) = (-1)^m (\ell - m)! \frac{Y_{\ell m}(\unit r)}{r^{\ell+1}} \qquad R_{\ell m} (\bm r) = (-1)^m \frac{r^\ell}{(\ell + m)!} Y_{\ell m}(\unit r).
\end{equation}

Assume that we can construct a sphere, centered at the center of mass of the central body, so that $\rho_M = 0$ outside the sphere and $\rho_m = 0$ inside the sphere. Then $r' > r$ in equation \ref{eqn:first-pe}, and Ref.~\cite{Gelderen1998TheSO} gives the identity
\begin{equation}
\frac{1}{|\bm r - \bm r'|} = \sum_{\ell, m} R_{\ell m}(\bm r) S_{\ell m}^*(\bm r'),
\label{eqn:1-over-r-expansion}
\end{equation}
where the sum is shorthand for $\sum_{\ell, m} = \sum_{\ell = 0}^\infty \sum_{m=-\ell}^\ell$.

Suppose the vector $\bm{D}$ points from the center of mass of the central body to the center of mass of the asteroid. Then we may write $ \bm r' = \bm D + \bm u$, and Ref.~\cite{Gelderen1998TheSO} gives
\begin{equation}
S_{\ell m}(\bm r') = \sum_{\ell', m'} (-1)^{\ell'}R^*_{\ell' m'}(\bm u)S_{\ell+\ell', m + m'} (\bm D).
\end{equation}

Let us define complex, unitless constants
\begin{equation}
J_{\ell m} = \int_\mathcal{P} d^3 r \rho_M(\bm r) \frac{R_{\ell m}(\bm r)}{\mu_M a_M^\ell}
\qquad
K_{\ell m} = \int_\mathcal{P} d^3 r \rho_m(\bm r) \frac{R_{\ell m}(\bm r)}{\mu_m a_m^\ell}
\label{eqn:jlm}
\end{equation}
where the integrals are taken in the inertial and the body-fixed frames respectively. The symmetry relations of $R_{\ell m}$ guarantee $J_{\ell m} = (-1)^m J_{\ell, -m}$ and similarly for $K_{\ell m}$. We have chosen $\mu_M$, $\mu_m$ to be the masses of the central body and asteroid respectively, and $a_M$ and $a_m$ to be scale lengths of both bodies (defined more precisely later). Then, by rotating $K_{\ell m}$ from the body-fixed axis orientation to the inertial axis orientation, we may use $J_{\ell m}$ and $K_{\ell m}$ to simplify the potential energy. Rotations of the spherical harmonics are given by the Wigner-D matrices, via the formula
\begin{equation}
Y_{\ell m}(M\bm r) = \sum_{m'=-\ell}^\ell \mathcal{D}^\ell_{mm'}(M)^* Y_{\ell m'}(\bm r).
\end{equation}
The corrolary of this formula for the regular solid spherical harmonic is
\begin{equation}
R_{\ell m} (M \bm r) = \sum_{m'=-\ell}^\ell (-1)^{m+m'} \frac{(\ell + m')!}{(\ell + m)!}\mathcal{D}^\ell_{mm'}(M)^* R_{\ell m'} (\bm r)
\end{equation}
where $\mathcal{D}^\ell_{mm'}$ are the Wigner-D matrix elements. In practice, these are computed from the Euler angles corresponding to the rotation $M$, which we specified in section \ref{sec:coordinates} are $z-y-z$ Euler angles. With this choice of axes, the Wigner-D matrix elements are real with matrix elements given by
\begin{equation}
\begin{aligned}
d^\ell_{mm'}(\beta) = &\sqrt{(\ell+m)!(\ell-m)!(\ell+m')!(\ell-m')!}\\
& \sum_{s=s_\text{min}}^{s_\text{max}} \brackets{\frac{(-1)^{m-m'+s}\parens{\cos\frac{\beta}{2}}^{2\ell + m' - m - 2s}\parens{\sin \frac{\beta}{2}}^{m-m'+2s}}{(\ell+m'-s)!(\ell-m-s)!(m-m'+s)!s!}}
\end{aligned}
\label{eqn:wigner-d-elements}
\end{equation}
where $s_\text{min} = \max(0, m'-m)$ and $s_\text{max} = \min(\ell+m', \ell-m)$,
and $\mathcal{D}^\ell_{mm'}(\alpha, \beta, \gamma) = d^j_{mm'}e^{-im\alpha} e^{-im'\gamma}$. For neatness, we define the unitless constants
\begin{equation}
C_{\ell \ell' m m' m''}(\alpha, \beta, \gamma, t) = D(t)(-1)^{\ell'+m'+m''} S^*_{\ell+\ell', m + m'} (\bm D (t)) a_M^\ell a_m^{\ell'} \frac{(\ell' + m'')!}{(\ell' + m')!} \mathcal{D}(\bm q)^\ell_{m'm''}(M)^*
\label{eqn:c}
\end{equation}
where $\bm D(t)$ is purely dependent on time because the orbital path of the asteroid is assumed to be Keplerian. Then, combining equations \ref{eqn:1-over-r-expansion} through \ref{eqn:wigner-d-elements}, we get a potential energy of
\begin{equation}
V = -\frac{G \mu_M \mu_m}{D(t)}\sum_{\ell, m}\sum_{\ell', m'} \sum_{m''=-\ell'}^{\ell'} C_{\ell \ell' m m' m''}(\alpha, \beta, \gamma, t) J_{\ell, m} K_{\ell' m''}.
\label{eqn:pe}
\end{equation}

\subsection{Kinetic energy}
The inertia tensor of the asteroid is expressed as a matrix $I$. Written in terms of spherical harmonics in the body-fixed frame, we have
\begin{equation}
\begin{aligned}
I = &\int_\mathcal{A} d^3 r\rho_m(\bm r) \parens{\begin{matrix}
y^2 + z^2 & -xy & -xz \\
-xy & x^2 + z^2 & -xz \\
-xz & -yz & x^2 + y^2 \\
\end{matrix}}\\
= &\frac{2}{3}\mu_m a_m^2\parens{\begin{matrix}
K_{20}-3K_{2,-2}-3K_{22}+1 & 3i \parens{K_{22}-K_{2,-2}} & \frac{3}{2}\parens{K_{21}-K_{2,-1}}\\
3i \parens{K_{22}-K_{2,-2}} & K_{20}+3K_{2,-2}+3K_{22}+1 & -\frac{3}{2}i\parens{K_{21}-K_{2,-1}}\\
\frac{3}{2}\parens{K_{21}-K_{2,-1}} & -\frac{3}{2}i\parens{K_{21}-K_{2,-1}} & -2K_{20}+1\\
\end{matrix}}
\end{aligned}
\label{eqn:moi}
\end{equation}
where we have defined, for neatness,
\begin{equation}
a_m^2=\frac{1}{\mu_m}\int_\mathcal{A} d^3 r \rho_m(\bm r) r^2.
\label{eqn:am}
\end{equation}
With this definition of $a_m$, and an arbitrary definition of $a_M$ which may be taken to be similar to equation \ref{eqn:am} or a more convenient unit such as the central body radius, the parameters $J_{\ell m}$ and $K_{\ell m}$ of equation \ref{eqn:jlm} are completely defined. Note that $a_m$ is similar to the radius of the asteroid; if the asteroid is a sphere of uniform density and radius $r$, then $a_m = r\sqrt{3/5}$.

@@

To calculate the kinetic energy, we write the angular velocity in the body-fixed frame in terms of the Euler angles:
\begin{equation}
\bm \omega_{xyz} = \parens{\begin{matrix}
0 & -\sin \alpha & \sin \beta \cos \alpha \\
0 & \cos \alpha & \sin \beta \sin \alpha \\
1 & 0 & \cos \beta
\end{matrix}}
\parens{\begin{matrix}
\dot \alpha\\
\dot \beta \\
\dot \gamma
\end{matrix}}.
\label{eqn:ang-vel}
\end{equation}
The rotational kinetic energy is then
\begin{equation}
K = \frac{1}{2} \bm\omega^T I \bm \omega
\label{eqn:ke}
\end{equation}
where $\bm \omega$ and $I$ are both given in the body-fixed frame.

\subsection{Hamiltonian and Equations of motion}
As a simplifying assumption, we will not model the effect of the central body's non-sphericity on the asteroid orbit. This effect has been studied in much detail and is small. We therefore separate the position dynamical variables from the angular variables and derive the standard equations of motion for asteroid position:
\begin{equation}
\dot{\bm{v}} = -\frac{GM}{r^3}\bm{r}\qquad \dot{\bm{r}} = \bm{v}.
\end{equation}
The position variables will no longer be discussed.

The Lagrangian for the rotating part of the system is $L=T-V$, with kinetic and potential energy given by equations \ref{eqn:ke} and \ref{eqn:pe} respectively. Defining $\bm q = \parens{\alpha, \beta, \gamma}^T$, we can write equation \ref{eqn:ang-vel} as $\bm \omega_{xyz} = A\bm {\dot q}$, in which case
\begin{equation}
L = \frac{1}{2}\bm {\dot q}^T A^T I A \bm {\dot q} - V_0\sum_{\ell, m}\sum_{\ell', m'} \sum_{m''=-\ell'}^{\ell'} C_{\ell \ell' m m' m''}(\bm q, t) J_{\ell, m} K_{\ell' m''}.
\end{equation}
where $V_0= -G \mu_M \mu_m / D(t)$ is the Newtonian potential energy of point masses $\mu_M$ and $\mu_m$. This corresponds to a Hamiltonian of
\begin{equation}
H = \frac{1}{2}\bm p^T (A^{-1}(\bm q)) I^{-1} (A^{-1}(\bm q))^T \bm p + V_0\sum_{\ell, m}\sum_{\ell', m'} \sum_{m''=-\ell'}^{\ell'} C_{\ell \ell' m m' m''}(\bm q, t) J_{\ell, m} K_{\ell' m''}
\label{eqn:hamiltonian}
\end{equation}
with the conjugate momenta defined as
\begin{equation}
\bm p = A^T I A \bm {\dot q} = A^T I \bm \omega_{xyz}
\qquad
\bm \omega_{xyz} = A \bm {\dot q} = I^{-1}(A^{-1})^T \bm p.
\label{eqn:momenta}
\end{equation}
The inverse of $A$ is computed as
\begin{equation}
A^{-1} = \frac{1}{\sin{\beta}} \parens{\begin{matrix}
-\cos\alpha \cos \beta & -\sin \alpha \cos \beta & \sin \beta \\
-\sin \alpha \sin \beta & \cos \alpha \sin \beta & 0 \\
\cos \alpha & \sin \alpha & 0
\end{matrix}}.
\label{eqn:a-inv}
\end{equation}
The equations of motion of the system are then a system of six first-order ordinary differential equations:
\begin{equation}
\bm{\dot q} = \nabla_{\bm{p}} H \qquad \bm{\dot p} = -\nabla_{\bm{q}} H.
\label{eqn:first-eom}
\end{equation}
We compute these derivatives manually:
\begin{equation}
\bm{\dot q} = A^{-1}I^{-1}(A^{-1})^T\bm p
\label{eqn:eom-q}
\end{equation}
and
\begin{equation}
\dot p_i = \bm p^T \frac{\partial A^{-1}}{\partial q_i} I^{-1} (A^{-1})^T \bm p -V_0 \sum_{\ell, m}\sum_{\ell', m'} \sum_{m''=-\ell'}^{\ell'} J_{\ell, m} K_{\ell' m''} \frac{\partial}{\partial q_i}C_{\ell \ell' m m' m''}(\bm q, t).
\label{eqn:eom-p}
\end{equation}
Explicitly, the derivatives of $A^{-1}$ are
\begin{equation}
\begin{aligned}
\frac{\partial A^{-1}}{\partial \alpha} = & \frac{1}{\sin \beta}\parens{\begin{matrix}
\sin \alpha \cos \beta & -\cos \alpha \cos \beta & 0\\
-\cos \alpha \sin \beta & -\sin \alpha \sin \beta & 0 \\
-\sin \alpha & \cos \alpha & 0\\
\end{matrix}}\\
\frac{\partial A^{-1}}{\partial \beta} = & \frac{1}{\sin^2\beta}\parens{\begin{matrix}
\cos \alpha & \sin \alpha & 0\\
0 & 0 & 0\\
-\cos\alpha \cos\beta & -\sin \alpha \cos \beta & 0\\
\end{matrix}}\qquad
\frac{\partial A^{-1}}{\partial \gamma} = 0.
\end{aligned}
\end{equation}
The $C$ coefficients depend on $\bm q$ through the Wigner-D matrices, which obey
\begin{equation}
\nabla_{\bm q}\mathcal{D}^\ell_{m m'}(\bm q) = \parens{\begin{matrix}
-im\alpha \mathcal{D}^\ell_{m m'}(\bm q)\\
e^{-im\alpha} e^{-im'\gamma}\frac{\partial}{\partial \beta} d^j_{mm'}\\
-im'\gamma \mathcal{D}^\ell_{m m'}(\bm q)\\
\end{matrix}}.
\label{eqn:wigner-d-derivatives}
\end{equation}
Finally, the derivative of the matrix elements $d^j_{mm'}$ are given by multiplying each term in the sum over $s$ of equation \ref{eqn:wigner-d-elements} by
$$\frac{1}{2}\parens{(m-m'+2s) \cot \frac{\beta}{2} - (2\ell + m' - m - 2s) \tan \frac{\beta}{2}},$$
except when $\beta = 0$ or $\pi$. In these cases, the derivative is zero except when $\beta=0$ and $m-m'+2s=1$, or $\beta = \pi$ and $2 \ell + m' - m - 2s=1$, in which case the derivative of the trigonometric functions is $1/2$.

The dynamical variables $\bm q$, $\bm p$ can then be integrated numerically to solve for the full dynamics of the system.

>>>>

\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{bbold}
\usepackage{siunitx}

\newcommand{\unit}[1]{\hat{\mathbf{#1}}}
\newcommand{\parens}[1]{\left( #1 \right)}
\newcommand{\brackets}[1]{\left[ #1 \right]}
\newcommand{\jtd}[1]{{\color{red}\textit{#1}}}
\newcommand{\quat}[1]{\widetilde{\bm{#1}}}
\usetikzlibrary{arrows.meta}

@@

@@

We define a rotation matrix $M$ such that $M\bm{r} = \bm{R}$, where $\bm{r}$ is in the body-fixed frame and $\bm{R}$ is in the inertial frame. It will be useful for us to represent this matrix in two different ways; one is with a normalized quaternion $\quat q$, and the other is with $z-y-z$ Euler angles named $\alpha$, $\beta$, and $\gamma$. In particular, we define
\begin{equation}
M = R_z(\alpha) R_y(\beta) R_z(\gamma)
\label{eqn:euler-angles}
\end{equation}
where $R_i(\theta)$ is a rotation around the $i$th unit vector by $\theta$ (figure \ref{fig:euler-angles}).

@@

We use the quaternion $\quat q$ as our dynamical variable because Euler angles suffer from gimbal lock. For the sake of this paper, we write quaternions as $\quat q = r + i \bm i + j \bm j + k \bm k$, with the complex conjugate written as $\quat q^* = r - i \bm i - j \bm j - k \bm k$. The inverse is then $\quat q^{-1} = \quat q^* / (\quat q \quat q^*)$, and we define real and imaginary parts as $\Re \quat q = r$; $\Im \quat q = i \unit x + j \unit y + k \unit z$. See a description of quaternions in Ref.~\cite{2008arXiv0811.2889G}

Normalized quaternions express rotations of an angle $\theta$ by $\quat q = \cos\frac{\theta}{2} + \sin\frac{\theta}{2}\quat n$ where $\quat n = 0 + n_x \bm i + n_y \bm j + n_z\bm k$. This rotation may be applied to any vector $\bm r$ by
\begin{equation}
\bm R = \Im \brackets{\quat q^{-1} \quat r\quat q }.
\label{eqn:quat-rot}
\end{equation}
Due to the real part of $\quat q$ which does not vanish for $\theta \rightarrow 0$, quaternions do not suffer from gimbal lock. Using the above notation, we can give the full equation of quaternion multiplication:
\begin{equation}
(q_0 + \quat q) (p_0 + \quat p) = q_0 p_0 - \bm p \cdot \bm q + p_0 \bm q + q_0 \bm p + \bm q \times \bm p.
\label{eqn:quat-mult}
\end{equation}
Note that quaternion multiplication is not commutative.

We convert from Euler angles to quaternions by writing equation \ref{eqn:euler-angles} in terms of quaternions:
\begin{equation}
\quat q = \parens{\begin{matrix} \cos \frac{\alpha}{2} \\ 0\\ 0 \\\sin \frac{\alpha}{2}\end{matrix}}
\parens{\begin{matrix} \cos \frac{\beta}{2} \\ 0 \\\sin \frac{\beta}{2}\\ 0\end{matrix}}
\parens{\begin{matrix} \cos \frac{\gamma}{2} \\ 0\\ 0 \\\sin \frac{\gamma}{2}\end{matrix}} = \parens{\begin{matrix}
\frac{\alpha}{2} \cos \frac{\beta}{2} \cos  \cos \frac{\gamma}{2} - \frac{\alpha}{2} \cos \frac{\beta}{2}\sin \sin \frac{\gamma}{2}\\
\frac{\alpha}{2} \sin \frac{\beta}{2} \cos  \sin \frac{\gamma}{2} - \frac{\alpha}{2} \sin \frac{\beta}{2}\sin \cos \frac{\gamma}{2}\\
\frac{\alpha}{2} \sin \frac{\beta}{2} \cos  \cos \frac{\gamma}{2} + \frac{\alpha}{2} \sin \frac{\beta}{2}\sin \sin \frac{\gamma}{2}\\
\frac{\alpha}{2} \cos \frac{\beta}{2} \cos  \sin \frac{\gamma}{2} + \frac{\alpha}{2} \cos \frac{\beta}{2}\sin \cos \frac{\gamma}{2}\\
\end{matrix}}.
\label{eqn:euler-to-quat}
\end{equation}
This equation can be reversed for $\alpha$, $\beta$, $\gamma$ as follows:
\begin{equation}
\tan \alpha = \frac{jk - ir}{jr + ik}\qquad \cos \beta = 1 - 2(i^2 + j^2) \qquad \tan \gamma = \frac{jk + ir}{jr - ik} .
\label{eqn:quat-to-euler}
\end{equation}

\subsection{Tidal Torque}
We start by expressing the gravitational potential energy of the central body in its most general form:
\begin{equation}
V(\bm R') = -G\int_\mathcal{P} d^3 R \rho_M(\bm R) \frac{1}{|\bm{R}-\bm{R'}|}.
\label{eqn:first-pe}
\end{equation}
This integral is computed in the inertial frame. We represent the central body density as $\rho_M$ and its volume as $\mathcal{P}$. To reduce equation \ref{eqn:first-pe}, we define the unnormalized spherical harmonics $Y_{\ell m}(\theta, \phi) = P_{\ell m}(\cos \theta)e^{im\phi}$, where $P_{\ell m}$ are the associated Legendre Polynomials without the Condon-Shortley phase. The regular and irregular spherical harmonics are then defined as
\begin{equation}
S_{\ell m}(\bm r) = (-1)^m (\ell - m)! \frac{Y_{\ell m}(\unit r)}{r^{\ell+1}} \qquad R_{\ell m} (\bm r) = (-1)^m \frac{r^\ell}{(\ell + m)!} Y_{\ell m}(\unit r).
\end{equation}

Assume that we can construct a sphere, centered at the center of mass of the central body, so that $\rho_M = 0$ outside the sphere and $\rho_m = 0$ inside the sphere. Then $R' > R$ in equation \ref{eqn:first-pe}, and Ref.~\cite{Gelderen1998TheSO} gives the identity
\begin{equation}
\frac{1}{|\bm R - \bm R'|} = \sum_{\ell, m} R_{\ell m}(\bm R) S_{\ell m}^*(\bm R'),
\label{eqn:1-over-r-expansion}
\end{equation}
where the sum is shorthand for $\sum_{\ell, m} = \sum_{\ell = 0}^\infty \sum_{m=-\ell}^\ell$.

Suppose the vector $\bm{D}$ points from the center of mass of the central body to the center of mass of the asteroid, again in the inertial frame. Then we may write $ \bm R' = \bm D + \bm R$, and Ref.~\cite{Gelderen1998TheSO} gives
\begin{equation}
S_{\ell m}(\bm R') = \sum_{\ell', m'} (-1)^{\ell'}R^*_{\ell' m'}(\bm U)S_{\ell+\ell', m + m'} (\bm D).
\end{equation}

Define the complex, unitless constants
\begin{equation}
J_{\ell m} = \int_\mathcal{P} d^3 R \rho_M(\bm R) \frac{R_{\ell m}(\bm R)}{\mu_M a_M^\ell}
\qquad
K_{\ell m} = \int_\mathcal{A} d^3 r \rho_m(\bm r) \frac{R_{\ell m}(\bm r)}{\mu_m a_m^\ell}
\label{eqn:jlm}
\end{equation}
where the integrals are taken in the inertial and the body-fixed frames respectively. Here, $\rho_m$ denotes the density of the asteroid and $\mathcal{A}$ denotes its volume. The symmetry relations of $R_{\ell m}$ guarantee $J_{\ell m} = (-1)^m J_{\ell, -m}$ and similarly for $K_{\ell m}$. We have chosen $\mu_M$, $\mu_m$ to be the masses of the central body and asteroid respectively, and $a_M$ and $a_m$ to be scale lengths of both bodies (defined more precisely later). Then we can express the potential as
\begin{equation}
V(\bm U) = -G \brackets{\sum_{\ell, m} \mu_M a_M^\ell J_{\ell m} \sum_{\ell', m'} (-1)^{\ell'}R_{\ell' m'}(\bm U)S^*_{\ell+\ell', m + m'} (\bm D)}.
\label{eqn:pe-unrotated}
\end{equation}
To make use of $K_{\ell m}$ in equation \ref{eqn:jlm}, we need to rotate $\bm U$ from the inertial frame into $\bm u$ in the body fixed frame via $\bm U = M \bm u$. The rotation of a spherical harmonic satisfies
\begin{equation}
Y_{\ell m}(M\bm u) = \sum_{m'=-\ell}^\ell (-1)^{m+m'}\sqrt{\frac{(\ell-m')!(\ell+m)!}{(\ell+m')!(\ell-m)!}} \mathcal{D}^\ell_{mm'}(M)^* Y_{\ell m'}(\bm u).
\end{equation}
where $\mathcal{D}$ are the Wigner-$D$ matrices. The corollary of this formula for the regular solid spherical harmonic is
\begin{equation}
R_{\ell m} (M \bm u) = \sum_{m'=-\ell}^\ell \sqrt{\frac{(\ell-m')!(\ell+m')!}{(\ell-m)!(\ell+m)!}} \mathcal{D}^\ell_{mm'}(M)^* R_{\ell m'} (\bm u)
\label{eqn:rlm-rotations}
\end{equation}
where $\mathcal{D}^\ell_{mm'}$ are the Wigner-D matrix elements. In practice, these are computed from the Euler angles corresponding to the rotation $M$, which we specified in section \ref{sec:coordinates} are $z-y-z$ Euler angles. With this choice of Euler angles, the Wigner-D matrix elements are real with matrix elements given by
\begin{equation}
\begin{aligned}
d^\ell_{mm'}(\beta) = &\sqrt{(\ell+m)!(\ell-m)!(\ell+m')!(\ell-m')!}\\
& \sum_{s=s_\text{min}}^{s_\text{max}} \brackets{\frac{(-1)^{m-m'+s}\parens{\cos\frac{\beta}{2}}^{2\ell + m' - m - 2s}\parens{\sin \frac{\beta}{2}}^{m-m'+2s}}{(\ell+m'-s)!(\ell-m-s)!(m-m'+s)!s!}}
\end{aligned}
\label{eqn:wigner-d-elements}
\end{equation}
where $s_\text{min} = \max(0, m'-m)$ and $s_\text{max} = \min(\ell+m', \ell-m)$,
and $\mathcal{D}^\ell_{mm'}(\alpha, \beta, \gamma) = d^j_{mm'}e^{-im\alpha} e^{-im'\gamma}$.

Combining equations \ref{eqn:rlm-rotations} and \ref{eqn:pe-unrotated},
\begin{equation}
V(\bm u) = -G \brackets{\sum_{\ell, m} \mu_M a_M^\ell J_{\ell m} \sum_{\ell', m'} (-1)^{\ell'}S^*_{\ell+\ell', m + m'} (\bm D)\sum_{m''=-\ell'}^{\ell'} \sqrt{\frac{(\ell'-m'')!(\ell'+m'')!}{(\ell'-m')!(\ell'+m')!}} \mathcal{D}^{\ell'}_{m'm''}(M)^* R_{\ell' m''} (\bm u)}.
\end{equation}

The tidal torque applied to an asteroid at a point $\bm u$ is given by $d\bm \tau(\bm u) = -\rho_m(\bm u) d^3u (\bm u \times (\nabla V_{\bm u}(\bm u)))$ in the body-fixed frame, which leaves
\begin{equation}
\begin{aligned}
\bm\tau = &G\int_\mathcal{A}d^3u \rho_m(\bm u)  \bm u \times \nabla \left[\sum_{\ell, m} \mu_M a_M^\ell J_{\ell m} \sum_{\ell', m'} (-1)^{\ell'}S^*_{\ell+\ell', m + m'} (\bm D)
\right.\\
& \left. \sum_{m''=-\ell'}^{\ell'} \sqrt{\frac{(\ell'-m'')!(\ell'+m'')!}{(\ell'-m')!(\ell'-m')!}} \mathcal{D}^{\ell'}_{m'm''}(M)^* R_{\ell' m''} (\bm u)\right].
\end{aligned}
\label{eqn:first-tidal-torque}
\end{equation}
Using the identity
\begin{equation}
\begin{aligned}
\bm u \times \nabla R_{\ell m}(\bm u)=&\frac{1}{2}[(i\unit x - \unit y)(\ell-m+1)R_{\ell,m-1}(\bm u)\\
&+(i\unit x+\unit y)(\ell+m+1)R_{\ell,m+1}(\bm u)+2im\unit z R_{\ell m}(\bm u)],
\end{aligned}
\end{equation}
which was derived from Ref.~\cite{Gelderen1998TheSO}'s analysis of the angular momentum operators applied to spherical harmonics, we write equation \ref{eqn:first-tidal-torque} as
\begin{equation}
\begin{aligned}
\bm \tau = &G\frac{\mu_m\mu_M}{2}\left[\sum_{\ell, m} a_M^\ell J_{\ell m} \sum_{\ell', m'}a_m^{\ell'}S^*_{\ell+\ell', m + m'} (\bm D) (-1)^{\ell'}\sum_{m''=-\ell'}^{\ell'} \sqrt{\frac{(\ell'-m'')!(\ell'+m'')!}{(\ell'-m')!(\ell'+m')!}}  \mathcal{D}^{\ell'}_{m'm''}(M)^* \right. \\
&\bigg[(i\unit x - \unit y)(\ell'-m''+1)K_{\ell',m''-1} +(i\unit x+\unit y)(\ell'+m''+1)K_{\ell',m''+1}+2im''\unit z K_{\ell'm''}\bigg] \Bigg].
\label{eqn:tidal-torque}
\end{aligned}
\end{equation}
This equation contains complex values, but it was numerically confirmed that $\bm \tau$ is a real-valued vector.

Note that each sum in this equation is weighted by a factor of $a_M^\ell a_m^{\ell'}/D^{\ell+\ell'}$. Since $D \gg a_M \gg a_m$ in general, the leading order torques come from low $\ell$ and $\ell'$. However, $\ell'$ contributes nothing because the inner bracketed term is zero, and $K_{\ell' = 1} = 0$ if we choose the origin of the body-fixed axes to be the center of mass of the asteroid. The leading order of equation \ref{eqn:tidal-torque} is therefore $1/D^3$ with $\ell = 0$ and $\ell' = 2$.

\subsection{Moment of Inertia}
The inertia tensor of the asteroid is expressed as a matrix $I$. Written in terms of spherical harmonics in the body-fixed frame, we have
\begin{equation}
\begin{aligned}
I = &\int_\mathcal{A} d^3 r\rho_m(\bm r) \parens{\begin{matrix}
y^2 + z^2 & -xy & -xz \\
-xy & x^2 + z^2 & -xz \\
-xz & -yz & x^2 + y^2 \\
\end{matrix}}\\
= &\frac{2}{3}\mu_m a_m^2\parens{\begin{matrix}
K_{20}-3K_{2,-2}-3K_{22}+1 & 3i \parens{K_{22}-K_{2,-2}} & \frac{3}{2}\parens{K_{21}-K_{2,-1}}\\
3i \parens{K_{22}-K_{2,-2}} & K_{20}+3K_{2,-2}+3K_{22}+1 & -\frac{3}{2}i\parens{K_{21}-K_{2,-1}}\\
\frac{3}{2}\parens{K_{21}-K_{2,-1}} & -\frac{3}{2}i\parens{K_{21}-K_{2,-1}} & -2K_{20}+1\\
\end{matrix}}
\end{aligned}
\label{eqn:moi}
\end{equation}
where we have defined, for neatness,
\begin{equation}
a_m^2=\frac{1}{\mu_m}\int_\mathcal{A} d^3 r \rho_m(\bm r) r^2.
\label{eqn:am}
\end{equation}
With this definition of $a_m$, and an arbitrary definition of $a_M$ which may be taken to be similar to equation \ref{eqn:am} or a more convenient unit such as the central body radius, the parameters $J_{\ell m}$ and $K_{\ell m}$ of equation \ref{eqn:jlm} are completely defined. Note that $a_m$ is similar to the radius of the asteroid; if the asteroid is a sphere of uniform density and radius $r$, then $a_m = r\sqrt{3/5}$.

@@

\subsection{Equation of Motion}
As a simplifying assumption, we will not model the effect of the central body's non-sphericity on the asteroid orbit. This effect has been studied in much detail and is small. We therefore separate the position dynamical variables from the angular variables and derive the standard equations of motion for asteroid position in the inertial frame:
\begin{equation}
\dot{\bm{V}} = -\frac{G\mu_M}{r^3}\bm{D}\qquad \dot{\bm{D}} = \bm{V}.
\label{eqn:translaton-eom}
\end{equation}
The position variables will no longer be discussed.

Let $\bm\omega$ be the rotational velocity of the asteroid in the body-fixed frame. We have two equations of motion for our system; one governs the evolution of $\bm \omega$ and the other governs the evolution of $\bm q$. Starting with the latter, consider a fixed point $\bm r$ in the body-fixed frame on a rotating asteroid. The corresponding point $\bm R$ in the inertial frame satisfies $\dot{\bm R} = \Im[\quat q^{-1}(\bm \omega \times \bm r )\quat q]$. Furthermore, we can demand that $\bm \omega \perp \bm r$ such that $\bm \omega \times \bm r = \Im[\quat \omega \quat r]$ Plugging this into equation \ref{eqn:quat-rot}, we have
\begin{equation*}
\dot{\quat q^{-1}} \quat r \quat q + \quat q^{-1} \quat r \dot{\quat q} = \quat q^{-1}\quat \omega \quat r \quat q.
\end{equation*}
Substituting $\bm r \rightarrow \bm R$ and acknowledging that $\Re[\dot{\quat q} \quat q^{-1}]$,
\begin{equation*}
2\dot{\quat q^{-1}}\quat q\quat R = \quat q^{-1}\quat \omega \quat q \quat R.
\end{equation*}
Since $\bm R$ is arbitrary, this simplifies to
\begin{equation}
\dot{\quat q} = \frac{1}{2}\quat \omega\quat q.
\label{eqn:quat-eom}
\end{equation}

The second equation of motion are Euler's equations of motion given our expression for tidal torque. Namely,
\begin{equation}
\begin{aligned}
I_1 \dot \omega_1 - \omega_2 \omega_3 (I_2 - I_3) &= \tau_1\\
I_2 \dot \omega_2 - \omega_3 \omega_1 (I_3 - I_1) &= \tau_2\\
I_3 \dot \omega_3 - \omega_1 \omega_2 (I_1 - I_2) &= \tau_3\\
\end{aligned}
\label{eqn:omega-eom}
\end{equation}
The components of $\bm \tau$ are given by equation \ref{eqn:tidal-torque}, which was also in the body-fixed frame.

\subsection{Initial Values}
We assume that the asteroid begins the simulation in its minimum energy configuration; that is, with $\bm \omega \parallel \unit z$. The initial conditions are therefore that $\bm \omega = \omega \unit z$, and the quaternion is fixed to rotate $\hat z$ to $\bm \Omega$ in the inertial frame, which fixes Euler angles $\beta$ and $\gamma$. The third Euler angle $\alpha$ represents the initial roll of the asteroid, and therefore cannot be determined from physical data. However, it is an initial condition of the simulation. Once the Euler angles are known, the quaternion is computed via equation \ref{eqn:euler-to-quat}.

\subsection{Fit Parameters}
By requiring that the $\unit x, \unit y, \unit z$ axes coincide with principal axes, we force the diagonal components of $I$ to be zero. Thus, we may require $K_{21}=K_{2,-1}=0$ and $\Im K_{22}=\Im K_{2,-2}=0$ with no loss of generality. By requiring that the body-fixed frame be located at the center of mass of the asteroid, we also require $K_{1m}=0$. Finally, $K_{00}=1$ follows automatically from the definition. Thus, $K_{20}$, $\Re K_{22}$, and $K_{\ell>2, m}$ are the only physical parameters of the asteroid density distribution, beyond $a_m$.

As discussed before, the leading order of equation \ref{eqn:tidal-torque} is $\ell' = 2$. Therefore, the equation is proportional to $\mu_m a_m^2$. Factoring out this product results in an equation of motion independent of $\mu_m$, meaning that the asteroid dynamics are unaffected by the mass of the asteroid. This is expected, and it means that the density distribution of the asteroid can only be determined up to a scale factor.

Since the coefficient of $J_{lm}K_{l'm'}$ is proportional to $a_M^\ell a_m^{\ell'} / D^{\ell+\ell'}$, to first order in the asteroid density distribution, the system is also independent of $a_m$. When $K_{\ell' > 2}$ terms are included, $a_m$ scales the size of the higher order terms.
